---
title: "pivot-tables.Rmd"
author: "Garen T. Merritt"
date: "6/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# attach libraries

library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) #install.packages('kableExtra')

```

Intro to pivoting, group-by, summarize, and piping.

```{r}
# read in data

lobsters <- read_xlsx(here("data/lobsters.xlsx"), skip = 4)
```

# using skimr is great for reading non-numeric data as summary is   # good for numeric data

```{r}

# explore your data w/ skimr

skimr::skim(lobsters)

```

Group by, pipe, and summarize by one variable

```{r}
lobsters %>% group_by (year) %>% summarize(count_by_year = n())

```

Multiple variables group by

```{r}

lobsters %>% group_by(site, year) %>% summarize(count_by_year = n())

```

Summarize multiple variables, calc mean, std dev, and account for NAs and save data to variable for future use

```{r}

# lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear = n(), mean_size_mm = mean(size_mm))

# account for NAs

siteyear_summary <- lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear = n(), mean_size_mm = mean(size_mm, na.rm = TRUE), sd_size_mm = sd(size_mm, na.rm = TRUE))


```

Utilizing kable() to format tables

```{r}
# table w/new variable using kable()

siteyear_summary %>% kable()



```

# R code in-line in Markdown

"There are `r nrow(lobsters)` total lobsters in this report."
