---
title: "Creating graphs in R with 'ggplot2'"
author: "Garen T. Merritt"
date: "6/21/2022"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R and it's going to be fantastic.

```{r}
library(tidyverse)
library(readxl)
library(here)
```


```{r}
ca_np <- read_csv(here("data", "ca.np.csv"))
```


Creating different plots of varying colors, shapes, and sizes based on variable(s).

```{r}
 gg_base + geom_point(color = "salmon2", size = 3)
 gg_base +
 geom_point()
 gg_base +
 geom_point(aes(size = visitors, color = visitors), alpha = 0.5)
 gg_base + geom_col(aes(fill = visitors))
```

Using theme() to create major projects that you can manually tweak later

```{r}
gg_base + geom_point(aes(size = visitors, color = visitors), alpha = 0.5) + theme_minimal()
```


Multi-series ggplot graphs. Crucial lesson in grouping.

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) + geom_line()
```

Faceting ggplot graphs

```{r}
gg_np + geom_line(show.legend = FALSE) + theme_light() + labs(x = 'year', y = 'annual visitors') +          facet_wrap(~ park_name)
```

Export with ggsave(). You can adjust height, width, dpi, as well as export type.

```{r}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)
```

